---
title: "What makes Canadians rate their life: life feeling rating analysis"
author: "Man Fei, 1002129984"
date: "2020/12/21"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library("tidyverse")
library("ggpubr")
```

## Abstract
It is important to have a good feeling towards one's life. In this report, CES data is used to find out if happiness (Feelings_life in CES) for Canadian relates to their financial situation, education level, family situation(marriage, children), mental & physical health, and sex. The data is from the 2017 Canadian General Social Survey [1]. Multiple linear relationship between Canadians’ Feelings_life, and 10 different predictor variables are the interest for this report. To build more complex linear models, powers, logs and reciprocal transformation and AIC elimination are used to get a final model. Significance of the model is discussed, and in the next steps more methods can be used to get more complete model.

## Key words
Feeling life, Rating, Observational Study, Multiple Linear Regression, Satisfaction

## Introduction
One’s degree of satisfaction towards life is very important for people. It is essential because this greatly determines one's happiness and fulfillment in life. Satisfaction on life can be affected by many things. For example, one's economic standing affect his satisfaction towards life. If one spends more time on trying to make more money, there is less time foe him to spend time on what he likes. Another example could be marital status can affect one's feeling for life. Different marital status gives a person different sense of belonging, loneliness, or freedom etc, and all these feelings can affect people's feeling towards life.\

Multiple Linear Regression can be a good way to find out the correlation between feeling towards life and other aspects in life. It can model correlations between more than two variables, which is suitable for this circumstance [2]. Observational data can be put into MLR model to find out if there is multiple linear relationship [2], and further transformations of data can also be done.\

This report continue exploring CES data to find out if happiness (Feelings_life in CES) for Canadian relates to their financial situation, education level, family situation(marriage, children), mental & physical health, sex, and ages. The data is the dataset from the 2017 Canadian General Social Survey (Cycle 31, Family) [1]. To continue from a less-than-perfect report [3], further exploration is needed on whether there is a multiple linear relationship between Canadians’ Age, Feelings_life, age_at_first_birth, sex, marital_status, education, income_family, average_hours_worked, self_rated_health, and self_rated_mental_health, adding two more variables age_youngest_child_under_6 and total_children. In addition to the former project [3], a more significant model is built by more complex linear models. Box-Cox transformation, AIC backward elimination are implemented. Results show that some variable (age) is not as significant as others. The final model has a higher $R^2$ compared to former model, making it more significant in predicting life satisfaction feeling.\


```{r,echo=FALSE}
# Raw data from 2017 GSS data in file AAy2jjGy.csv,
# data cleaning done by gss_cleaning.R to get gss.csv file
dataAll<-read.csv("gss.csv",header = TRUE)
```

## Methodology
Code and data supporting this analysis is available at:  "https://github.com/Feymann7/STA304-final-project"
\

DATA\
Table 1: all data\
```{r,echo=FALSE}
## clean and get the 12 variables we want
dataSelected <- dataAll %>% select(age, 
                                   age_at_first_birth, 
                                   feelings_life, 
                                   sex, 
                                   marital_status,
                                   education,
                                   income_family,
                                   average_hours_worked,
                                   self_rated_health,
                                   self_rated_mental_health,
                                   age_youngest_child_under_6,
                                   total_children)%>% 
  # filter out missing data
  filter(!is.na(feelings_life)& feelings_life <= 10 & feelings_life >= 0)%>% 
  filter(!is.na(age)& age >= 0)%>%
  filter(!is.na(age_at_first_birth)& age_at_first_birth >= 0)%>% 
  filter(!is.na(sex))%>% 
  filter(!is.na(marital_status))%>% 
  filter(!is.na(education))%>% 
  filter(!is.na(income_family))%>% 
  filter(!is.na(average_hours_worked))%>% 
  filter(!is.na(self_rated_health))%>% 
  filter(!is.na(self_rated_mental_health))%>%
  filter(!is.na(age_youngest_child_under_6)& age_youngest_child_under_6 >= 0)%>%
  filter(!is.na(total_children)& total_children >= 0)
attach(dataSelected)
head(dataSelected, main = "Table 1: all data")
```
Table 1 characterizes the data I use from CES 2017 data set [2]. The first 6 available cases are include to show as examples. There are 12 interesting variables included in the data, and they are age, 
                                   age_at_first_birth, 
                                   feelings_life, 
                                   sex, 
                                   marital_status,
                                   education,
                                   income_family,
                                   average_hours_worked,
                                   self_rated_health,
                                   self_rated_mental_health,
                                   age_youngest_child_under_6,
                                   total_children. Five are numeric variables, age, age_at_first_birth, feelings_life, age_youngest_child_under_6, total_children. The other seven variables are categorical variables. By observing the data of those 12 variables, I find that feelings_life, which is the feeling towards life is quite related to the rest of variables. This triggers me to find if there is a linear relationship between them. Also, the right skewed normal curve of Feeling about life rating histogram below demonstrate feelings_life is not randomly distributedm but might be related to other things. I use Multiple Linear Model to try to find a model for it.\

```{r,echo=FALSE}
dataSelected %>% 
  ggplot(aes(x =feelings_life)) + 
  geom_histogram() +
  ggtitle("Feeling about life rating histogram") 
```

MODEL\

For my initial model 1, I run a multiple linear regression through all variables, using feelings_life as the response variable, and age, age_at_first_birth, feelings_life, sex, marital_status, education, income_family, average_hours_worked, self_rated_health, self_rated_mental_health, age_youngest_child_under_6, total_children as the predictors. The p-value for the global F-test is <2.2e-16, and the $R^2 = 0.3187$. This the a rather significant result for such large data set.\
```{r,echo=FALSE}
#Multiple Regression Model
model1<-lm(feelings_life ~age_at_first_birth 
           +age 
           +sex
           +marital_status
           +average_hours_worked
           +education
           +income_family
           +self_rated_health
           +self_rated_mental_health
           +age_youngest_child_under_6
           +total_children,
           data = dataSelected)
summary(model1)
```

```{r,echo=FALSE}
##Power transformations
library(alr4)
trans <- powerTransform(cbind(age_at_first_birth, age, total_children)~1)
summary(trans)

# transformed variables
tfeelings_life <- log(feelings_life)
tage <- log(age)
ttotal_children <- log(total_children)
model2 <- lm(feelings_life ~ age_at_first_birth + tage +sex+ average_hours_worked+
        marital_status+education+income_family+self_rated_health+self_rated_mental_health+
          age_youngest_child_under_6 + ttotal_children)
summary(model2)
```
However, I try to improve the model by Box-Cox Transformation. After I transform variables using Box-Cox transformation, I find out the $R^2$ is the same as model 1, but model 2 is more complicated in formula. So the original model 1 is retained.
I observe the data p-values and find out that although the global F-test is significant for model 1, some individual tests fail. For example, the p-value for self_rated_mental_healthPoor is insignificant while other categories for the same variable is significant. Based on this, I create a new model 3 using AIC elimination to solve this.\

```{r, echo=FALSE}
back1=step(model1, direction = "backward")
```

```{r, echo=FALSE}
model3<-lm(feelings_life ~age_at_first_birth  +sex+average_hours_worked+
        marital_status+education+income_family+self_rated_health+self_rated_mental_health+
        age_youngest_child_under_6+ total_children,
      data = dataSelected)
summary(model3)
```
\
After age is deleted from the linear model, Model 3 is the final model. The linear regression of model 3 is $y = \beta_{0} + \beta_{1}x_1 +\beta_{2}x_2 + \beta_{3}x_3 +\beta_{4}x_4 + \beta_{5}x_5 +\beta_{6}x_6 + \beta_{7}x_7 +\beta_{8}x_8 + \beta_{9}x_9 + \beta_{10}x_{10}$,
where y = feeling_life which is the response variable, $x_1$ - $x_9$ are predictor variables, 
$x_1$ = age_at_first_birth,  $x_2$ = sex, $x_3$ = average_hours_worked, $x_4$ = marital_status, $x_5$ = education, $x_6$ = income_family, $x_7$ = self_rated_health, $x_8$ = self_rated_mental health, $x_9$ = age_youngest_child_under_6, $x_{10}$ = total_children. The final model 3 has the smallest AIC and the simpliest predictor variable constitutions, for response variable feelings_life.\

# Results


# Discussion

SUMMARY

CONCLUSIONS

WEAKNESS & NEXT STEPS\

```{r}
par(mfrow=c(2,2))
plot(model3)
```

# References 
[1] General Social Survey (GSS), Cycle 31, 2017: Family. Retrieved December 18 2020, from http://www.chass.utoronto.ca
\
[2] Multiple Linear Regression. (2020). Retrieved December 18 2020, from http://www.stat.yale.edu/Courses/1997-98/101/linmult.htm
\

[3] What Makes Canadian Rate Their Feelings of Life. STA304, Winter 2020. Yiwen Feng, Yutong Yuan, Sirui Xu, Man Fei, University of Toronto.\


2. 5.3 - The Multiple Linear Regression Model | STAT 462. (2020). Retrieved December 18 2020, from https://online.stat.psu.edu/stat462/node/131/
3. gss.csv. (2020, December 21) Author :Rohan Alexander and Sam Caetano

4. Multiple Linear Regression. (2020). Retrieved December 18 2020, from http://www.stat.yale.edu/Courses/1997-98/101/linmult.htm